Rising from the Ashes II: Did the Birth of the Milky Way's Bar Give Birth to the Birth of the Abundance Bimodality

### 2024-10-07 paper/ms.tex
Update to match style of rest of text

\documentclass[twocolumn]{aastex631}

@@

\title{Rising from the Ashes II: Can Bar Formation Lead to Abundance Bimodalities?}

@@

\author{James W. Johnson}
\affiliation{Carnegie Science Observatories, Pasadena, CA, USA}

\author{Vadim Semenov}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Lars Hernquist}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Vedant Chandra}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\author{Charlie Conroy}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian, Cambridge, MA, USA}

\begin{abstract}
    The Milky Way is known to host at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation. Following a minor post-processing adjustment -— motivated by recent findings on the star formation efficiency in TNG -- which increases the \alphaFe{} values of older star particles, we demonstrate that this galaxy exhibits a pronounced bimodal distribution in the \alphaFe{}-\FeH{} plane. The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality is caused by AGN-induced quenching triggered by the formation of the Milky Way's bar. We argue that the post-processing step can be interpreted as the TNG model underproducing star formation in the densest regions at high-$z$.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of their abundances encodes the chemical history of a galaxy's gas phase. Two classes elements have received particular attention in the Milky Way due to their disparate formation channels: iron-peak (such as Fe) and $\alpha$-elements (produced through the $\alpha$-process, such as O and Mg). Iron-peak elements are produced in both Type~Ia and Type~II supernovae (SNe) whereas $\alpha$-elements are produced predominantly through Type~II SNe. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences shown in Figure~\ref{fig:fig1} \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}.

Broadly speaking, there are three approaches to explaining the bimodality. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. Third and finally, the bimodality is generated through a merger process, either by enhancing the star formation rate (SFR) of the Galaxy \citep{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}. Strong evidence that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports these scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}.

In \citet{2024arXiv240707985B} (hereafter Paper~I), we proposed an alternate scenario for the formation of the bimodality, driven by a brief ($\sim300\,Myr$) quiescent period in the Galaxy's history, which has two effects. First, the reduced SFR lowers the number of $\alpha$-producing Type~II SNe relative to Type~Ia\footnote{These SNe occur on different timescales (10s of Myr after star formation for Type~II, as compared to 100s of Myr to Gyrs for Type~Ia), so if the SFR drops, the number of Type~II SNe drops faster than Type~Ia.}, causing a sharp decline in the \alphaFe{} ratio in the star-forming gas. Second, the lower SFR results in fewer stars forming in the region between the high- and low-$\alpha$ sequences, reducing the occurance of this transitional population in present-day observations. This mechanism resembles two-phase infall models that incorporate a temporary halt in star formation \citep[][and references therein]{2024arXiv240511025S}, though the quiescent period in Paper~I is much shorter. The combination of these two effects creates the valley between the high- and low-$\alpha$ sequences.

In Paper~I, we used idealized simulations of a galaxy merger that triggered the quiescent period. However, we argued that the merger aspect of that work was not necessary. In this work, we study a subhalo, or galaxy, from the Illustris TNG50 simulation demonstrating this. This subhalo of interest (SoI) exhibits the sequence of events presented in Paper~I after a simple post-processing step which increases the \alphaFe{} of old star particles. This post-processing is motivated by recent work demonstrating that the star formation efficiency (SFE) of the dense gas present at high-$z$ is too low \citep{2024arXiv240909121H}. The SoI undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in Paper~I is possible in cosmological simulations and does not require a merger.

@@

\section{Methods}\label{sec:methods}
\subsection{IllustrisTNG Sample}\label{ssec:tng}
We use Illustris TNG50 \citep{2019MNRAS.490.3196P, 2019MNRAS.490.3234N}, a cosmological simulation of a $\sim50\,\textrm{cMpc}$ box at high resolution ($m_{\textrm{baryon}}\sim8.5\times10^4\,\Msun$). It uses the gravito-magneto-hydrodynamics code \texttt{AREPO} \citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}, along with the TNG model \citep{2013MNRAS.436.3031V, 2017MNRAS.465.3291W, 2018MNRAS.473.4077P}. This model includes several subgrid processes: a wind generation model, chemical enrichment from SNe and asymptotic giant branch stars, and thermal and kinetic feedback from AGN.

One piece of the TNG model of note for this work is the black hole (BH) accretion and feedback method \citep{2017MNRAS.465.3291W}. The BH accretion rate ($\dot{M}_{\textrm{BH}}$) is computed using the local structure of the gas phase with the Bondi-Hoyle-Lyttleton formula \citep{1939PCPS...35..405H,1944MNRAS.104..273B,1952MNRAS.112..195B}, with a maximum of the Eddington accretion rate ($\dot{M}_{\textrm{edd}}$). The model allows the BH to be either in a kinetic radio-mode or a thermal quasar-mode. If the Eddington ratio ($\dot{M}_{\textrm{BH}}/\dot{M}_{\textrm{edd}}$) exceeds a threshold ($M_{\textrm{BH}}$-dependent, but $\sim0.001$--$0.1$), the BH is in the quasar mode and injects a large amount of thermal energy into its surroundings.

Using the public catalog, we selected a sample of subhalos at $z=1.5$ (snapshot 40) according to the following criteria: (1) the subhalo is central (i.e., the most massive subhalo within its halo), and (2) the subhalo's stellar mass is between $10^{10}$ and $10^{10.5}\,\Msun/\textrm{h}$. There were a total of 168 subhalos that met both criteria. The chosen mass range is broadly consistent with the expected mass of the Milky Way at this redshift \citep{2013ApJ...771L..35V}. We choose to make our selection of galaxies at $z=1.5$ instead of at lower redshift because we wish to capture the \textit{formation} of multimodal structure. Mergers at lower redshift contribute very little to the Milky Way's disk stars \citep[e.g.,][]{2016ARA&A..54..529B}, and would act as a contaminant in our sample.

We examine the abundance distribution in the \MgFe{}-\FeH{} plane for this sample of subhalos by visual inspection. Few subhalos display multimodal structure, and when present, it is much weaker compared to that observed in the Milky Way. We then apply a post-processing step to the \MgFe{} of the stars, adding a value of $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$ to each star particle, where $t_{1.5}$ is the age of the universe at $z=1.5$ ($\sim4.3\,\Gyr$), and $t_{\textrm{form}}$ is the formation time of the star particle. After this adjustment, which we justify in Section~\ref{ssec:sfe} based on TNG underpredicting the SFE of dense gas, we observe that the abundance structure becomes significantly more pronounced.

@@

We select subhalo 172175 (the ID at snapshot 40) for its resemblance to the Milky Way. We then studied the main descendant of this subhalo at $z=0$ (subhalo 392276 at snapshot 99). We refer to this subhalo as our subhalo of interest (SoI). \red{Add some morphological details of gal here}

\subsection{Observations}\label{ssec:obs}
We make use of two observational data sets. First, we use the ASPCAP DR17 catalog of stellar abundances \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}. We make the same selection cuts as in Section~2.4 of Paper I. These criteria are meant to select giants with high quality abundance measurements as well as restricting the sample to only stars in the disk with angular momenta similar to the Sun's. This results in a sample of 54,777 stars. We use Fe to track total metallicity and Mg alone as an $\alpha$-element.

We then further considered a dataset of stellar ages from the APOKASC2 catalog \citep{2018ApJS..239...32P}. This catalog uses a combination of APOGEE spectroscopic parameters and \textit{Kepler} time series photometry to compute astroseismic ages. Using only stars with $25\%$ age uncertainties (taken as the maximum of the upper and lower uncertainty), we cross-match this catalog to our larger sample from ASPCAP which results in a sample of 1201 stars.

\subsection{One-Zone Chemical Evolution Model}\label{ssec:onezone_met}
James inserts his section

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276.pdf}
  \caption{\textbf{When old stars are $\alpha$-enhanced, our subhalo of interest from TNG displays a prominent bimodality.} The upper left panel shows the distribution in the \MgFe{}-\FeH{} plane of the Milky Way, demonstrating a clear bimodality (data selection given in text). The lower left panel shows the 1D histograms of \MgFe{} at fixed \FeH{} values of $-0.5$, $-0.25$, $0$, and $0.25$ (blue, orange, green, and red, respectively). In the Milky Way, the bimodality is strongest at low metallicities while disappearing at high metallicities. The middle column shows the same plots but for our TNG subhalo of interest (392276) and with the fixed \FeH{} values $0.25\,\dex$ lower. Only faint structure is seen in the lowest bin (blue, $-0.75\,\dex$). The right column shows the same subhalo but after increasing the \MgFe{} value of star particles formed before $z=1.5$ linearly with formation time (with a slope of $0.1\dex/\Gyr$). A clear bimodality is shown in these panels which, unlike in the Milky Way, is present at all metallicities.}
  \label{fig:fig1}
\end{figure*}

The main result of our paper is given in Figure~\ref{fig:fig1}. Here, we compare the abundance plane in the Milky Way (left column) to that of our subhalo of interest (middle and right columns). The upper panels show the 2D distribution in the space of \MgFe{}-\FeH{}. We have applied the standard \texttt{scipy} implementation of a gaussian kernel density estimator to a Cartesian grid of points. For each panel, we normalize so that the integral of the distribution is unity. Colors are plotted in a log scale ranging from $0.08$ to $15\,\dex^{-2}$. Contour lines are plotted at $0.1$, $1.5$, and $10\,\dex^{-2}$.

The colored vertical regions are indicated at $\FeH=-0.75$, $-0.5$, $-0.25$, and $0\,\dex$ in the Milky Way, and at bins $0.25\,\dex$ higher in the simulations. The lower panels show 1D histograms of \MgFe{} in bins centered on these values. The bins for the Milky Way/simulations have width $0.2$/$0.05\,\dex$. The rationale for the higher plotted \FeH{} in the simulations comes just from the empirical location of the bimodalities. The Milky Way shows a clear bimodal population, with a high-$\alpha$ sequence most clearly distinct from the low-$\alpha$ sequence at low metallicity. The two sequences merge around solar metallicity.

@@

The abundance distributions shown in Figure~\ref{fig:fig1} can be better understood by examining the evolution of \MgFe{} with time of the individual stars/star particles. In the upper panels of Figure~\ref{fig:alpha} we show the true distribution of \MgFe{} as a function of time for the fiducial SoI in the middle and for the post-processed, $\alpha$-enhanced subhalo to the right. We use age instead of formation time in order to better facilitate comparisons to observations. These panels show 2D histograms, with a logarithmic colormap normalized to the maximum of the plot. To prevent overlap, the values of \MgFe{} are given offsets of $0$, $-0.5$, $-1$, and $-1.5$, in order of increasing \FeH{}.

In the SoI, there is an age gap at $sim10.6\,\Gyr$, which we mark with a vertical dashed line in the upper row. Star particles older than this line have a much clearer gradient in time with \MgFe{} than stars that form after. In the $\FeH=-0.25$ bin, star particles which form directly after this line have a slightly reduced \MgFe{} than stars which form a short time later.

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $15\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC2 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is relatively minor, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

Interestingly, the Milky Way distribution (lower left panel) exhibits similarities to the $\alpha$-enhanced SoI. Specifically, in the $\FeH=-0.5$ and $-0.25$ bins (blue and orange, respectively), there appear to be two populations that overlap in age but remain distinct in \MgFe{}. These are the bins where the bimodality is most pronounced (as seen in the upper left panel of Figure~\ref{fig:fig1}).

>>>>

\documentclass[linenumbers, twocolumn]{aastex631}

@@

\title{Rising from the Ashes II: Did the Birth of the Milky Way's Bar Give Birth to the Birth of the Abundance Bimodality}

@@

\author{James W. Johnson}
\affiliation{The Observatories of the Carnegie Institution for Science, Pasadena, CA, USA}

\author{et al}

\begin{abstract}
    The Milky Way is known to host at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation. After a minor post-processing step which boosts the \alphaFe{} of old star particles, we demonstrate that this galaxy hosts a strongly bimodal distribution in the \alphaFe{}-\FeH{} plane. The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality is caused by AGN-induced quenching triggered by the formation of the Milky Way's bar. We argue that the post-processing step can be interpreted as the TNG model underproducing star formation in the densest regions at high-$z$.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of stellar surface abundances encodes the chemical history of a galaxy's gas phase. Two elements have received particular interest in the Milky Way: Fe and $\alpha$-elements (elements produced through the $\alpha$-process, such as O and Mg). Fe is produced in both Type~Ia and Type~II SNe whereas $\alpha$-elements are produced predominantly through Type~II SNe. Because these SNe occur on different timescales (10s of Myr after star formation for Type~II, as compared to 100s of Myr to Gyrs for Type~Ia), the ratio between $\alpha$-elements and Fe is expected to generally decrease with time.

Because of their separate formation channels, the two dimensional plane of \alphaFe{}-\FeH{} has received considerable interest. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}. This bimodal distribution is shown in the upper left panel of Figure~\ref{fig:fig1}.

There are essentially two approaches to explaining the bimodality. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}.

Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. Third and finally, the bimodality is generated through a merger process, either by enhancing the SFR of the Galaxy \citep{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}. The fact that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports these scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}.

In \citet{2024arXiv240707985B} we argued for an alternate formation scenario of the bimodality, for which we provide further support in this work. In this scenario, the bimodality is formed through a brief quiescent period in the Galaxy's history. During this period, which lasted $\sim300\,\Myr$ in that setup, two things happen. First, as the chemical evolution of the galaxy proceeds, fewer stars form. Second, because the SFR is lower, $\alpha$-element production drops and so the \alphaFe{} of the gas drops as well. This is conceptually similar to some variants of the two-phase infall scenarios which allow for a brief halt in the star formation rate \citep[][and references therein]{2024arXiv240511025S}, though the quiescent period is much shorter in \citet{2024arXiv240707985B}. The combination of these two effects is what leads to the valley in between the high- and low-$\alpha$ sequences.

In that work, we used idealized simulations of a galaxy merger that triggered a starburst which preceded the quiescent period. However, we argued that the merger aspect of that work was not necessary, but rather the quiescence. In this work, we study a subhalo from the Illustris TNG50 simulation which demonstrates this. This subhalo of interest (SoI) exhibits the sequence of events presented in \citet{2024arXiv240707985B} after a simple post-processing step which enhances the \alphaFe{} of old star particles. The SoI undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in \citet{2024arXiv240707985B} is possible in cosmological simulations and can occur in non-merger scenarios.

@@

\section{Methods}\label{sec:methods}
\subsection{IllustrisTNG Sample}\label{ssec:tng}
We have made use of the Illustris TNG50 simulation \citep{2019MNRAS.490.3196P, 2019MNRAS.490.3234N}, a cosmological simulation of a $\sim50\,\textrm{cMpc}$ box at high resolution ($m_{\textrm{baryon}}\sim8.5\times10^4\,\Msun$). It uses the gravito-magneto-hydrodynamics code \texttt{AREPO} \citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}, along with the TNG model \citep{2013MNRAS.436.3031V, 2017MNRAS.465.3291W, 2018MNRAS.473.4077P}. This model includes several subgrid processes: a wind generation model, chemical enrichment from supernovae and asymptotic giant branch stars, and thermal and kinetic feedback from AGN.

One piece of the TNG model of note for this work is the BH accretion and feedback method \citep{2017MNRAS.465.3291W}. The black hole accretion rate ($\dot{M_{\textrm{BH}}}$) is computed using the local structure of the gas phase with the Bondi-Hoyle-Lyttleton formula \citep{1939PCPS...35..405H,1944MNRAS.104..273B,1952MNRAS.112..195B}, with a maximum of the Eddington accretion rate ($\dot{M_{\textrm{edd}}}$). The model allows the BH to be either in a kinetic radio-mode or a thermal quasar-mode. If the Eddington ratio ($\dot{M_{\textrm{BH}}}/\dot{M_{\textrm{edd}}}$) exceeds a threshold ($M_{\textrm{BH}}$-dependent, but $\sim0.001--0.1$), the BH is in the quasar mode and injects a large amount of thermal energy into its surroundings.

Using the public catalog, we selected a sample of subhalos at $z=1.5$ (snapshot 40) according to the following criteria: (1) the subhalo is central (i.e., the most massive subhalo within its halo), and (2) the subhalo's stellar mass is between $10^{10}$ and $10^{10.5}\,\Msun/\textrm{h}$. There were a total of 168 subhalos that met both criteria. The chosen mass range is broadly consistent with the expected mass of the Milky Way at this redshift \citep{2013ApJ...771L..35V}. We chose to make our selection of galaxies at $z=1.5$ instead of at lower redshift because we wished to capture the \textit{formation} of any multimodal structure. We did not want contamination by mergers at lower redshift which we know contribute very little to the Milky Way's disk stars \citep[e.g.,][]{2016ARA&A..54..529B}.

We examined the abundance distribution in the \MgFe{}-\FeH{} plane of this sample of subhalos by eye. Not many subhalos had multimodal structure, and any structure that was present was relatively weak compared to that observed in the Milky Way. We then applied a post-processing to the \MgFe{} of stars by adding to each star particle a value of $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$, where $t_{1.5}$ is the age of the universe at $z=1.5$ ($\sim4.3\,\Gyr$) and $t_{\textrm{form}}$ is the formation time of each star particle. With this post-processing, we found that much more structure was generally present.

@@

We selected subhalo 172175 (the ID at snapshot 40) for its particular resemblance to the Milky Way. We then studied the main descendant of this subhalo at $z=0$ (subhalo 392276 at snapshot 99). We refer to this subhalo as our subhalo of interest (SoI).

\subsection{Bar Catalog}\label{ssec:barcat}
In order to study the evolution of the morphology of our SoI, we make use of the \citet{2022MNRAS.515.1524Z} catalog. They provide a decomposition of all galaxies with at least $10^4$ star particles, as well as an estimate of the bar strength. This taken to be the maximum of the second Fourier mode ($A_{2,\textrm{max}}$) within the bar region (not including spiral arms).

\subsection{Observations}\label{ssec:obs}
We make use of two observational data sets. First, we use the ASPCAP DR17 catalog of stellar abundances \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}. We make the same selection cuts as in \citet{2024arXiv240707985B}, described in their Section~2.4. These are meant to select only giants with high quality abundance measurements as well as restricting the sample to only stars in the disk with angular momenta similar to the Sun's. This results in a sample of 54,777 stars. We use Fe to track total metallicity and Mg alone as an $\alpha$-element.

We then further considered a dataset of stellar ages from the APOKASC2 catalog \citep{2018ApJS..239...32P}. This uses a combination of APOGEE spectroscopic parameters and \textit{Kepler} time series photometry to compute astroseismic ages. Using only stars with $25\%$ age uncertainties (taken as the maximum of the upper and lower uncertainty), we cross-match this catalog to our larger sample from ASPCAP which results in a sample of 1201 stars.

\subsection{One-Zone Chemical Evolution Model}\label{ssec:onezone_met}
One-zone galactic chemical evolution (GCE) models describe enrichment in an idealized gas reservoir in which newly synthesized metals mix instantaneously (see e.g. the reviews by \citealt{Tinsley1980} and \citealt{Matteucci2021}). Our parameter choices in this paper are based on the models in \citet{2022arXiv220402989C}. While they were interested in bursts of star formation at the transition between the stellar halo and thick disk formation epochs, we focus on smooth evolutionary histories in this paper. We integrate these models numerically using the publicly available {\tt Versatile Integrator for Chemical Evolution} ({\tt VICE}; \citealt{Johnson2020}).

The quantity that is of particular importance to our results is the star formation efficiency (SFE) timescale $\tau_\star$:
\begin{equation}
\tau_\star \equiv \frac{M_{\textrm{gas}}}{\dot{M}_{\textrm{star}}},
\end{equation}
where $M_{\textrm{gas}}$ and $\dot{M}_{\textrm{star}}$ are the ISM mass and star formation rate, respectively. This timescale can be equivalently expressed as the ratio of the corresponding surface densities $\Sigma_{\textrm{gas}}$ and $\dot{\Sigma}_{\textrm{star}}$. $\tau_\star$ is the inverse of the SFE itself, describing the rate of star formation per unit mass in the ISM. In principle, $\tau_\star$ should vary with $M_{\textrm{gas}}$ because the observed relation between $\dot{\Sigma}_{\textrm{star}}$ and $\Sigma_{\textrm{gas}}$ is non-linear; classically, $\dot{\Sigma}_{\textrm{star}} \propto \Sigma_{\textrm{gas}}^N$ where $N \approx 1.5$ (e.g. \citealt{Schmidt1959, Kennicutt1998}). For simplicity, we hold $\tau_\star$ constant in this paper, which corresponds to a linear relation. In section 3.5 below, we present models using $\tau_\star = 0.5$, $1$, and $5$ Gyr.

Following \citet{2022arXiv220402989C}, our models assume that there is initially no gas present in the galaxy (i.e. $M_{\textrm{gas}} = 0$ at $t = 0$). Zero metallicity gas accretes at a constant rate of $\dot{M}_\text{in} = 5\,\Msunyr$. The efficiency with which feedback sweeps up ISM material and ejects it to the CGM in an outflow is described by the mass loading factor $\eta \equiv \dot{M}_\text{out} / \dot{M}_\star$. We use $\eta = 2$ throughout this paper (i.e. for every solar mass of stars formed, $2\,\Msun$ of ISM gas is lost to an outflow). These parameter choices lead to an SFH that initially rises from zero and approaches a constant value of $\sim1.9\,\Msunyr$. With these choices, the predicted abundance evolution is insensitive to the overall normalization of the accretion and star formation histories, since a larger accretion rate forms more stars and metals but simultaneously introduces more H into the ISM.

We use the solar abundances of Fe and Mg from \citet{Magg2022} with an additional 0.04 dex to account for the effects of diffusion and gravitational settling (\citealt{Turcotte1998}; i.e. $\{Z_{\text{Fe},\odot}, Z_{\text{Mg},\odot}\} = \{13.7, 6.71\} \times 10^{-4}$). We adopt SN yields of Fe and Mg based on \citeauthor{Weinberg2023}'s \citeyearpar{Weinberg2023} recommendations for Type~II SNe:
\begin{itemize}

	\item $y_\text{Fe}^\text{II} = 4.73 \times 10^{-4}$

	\item $y_\text{Mg}^\text{II} = 6.52 \times 10^{-4}$

	\item $y_\text{Fe}^\text{Ia} = 12 \times 10^{-4}$

	\item $y_\text{Mg}^\text{Ia} = 0$,

\end{itemize}
where the subscripts and superscripts denote the element and SN type, respectively. We select this value of $y_\text{Fe}^\text{Ia}$ as it corresponds to $\sim30\%$ ($\sim70\%$) of Fe arising from Type~II (Ia) SNe at solar \FeH{}. These yields are population-averaged, meaning that they quantify the amount of metal production per unit mass of star formation under the given enrichment channel (e.g. a hypothetical $1000\,\Msun$ stellar population would produce a total of $0.652\,\Msun$ of Mg through Type~II SNe). {\tt VICE} models Type~II SNe as exploding instantaneously following an episode of star formation. For Type~Ia SNe, we assume that metals are produced according to the empirically motivated $t^{-1.1}$ power-law based volumetric SN rates as a function of redshift and the cosmic SFH \citep[e.g.][]{Maoz2012}.

We use the yields recommended by \citet{Weinberg2023} because they are based on an empirical determination of the mean $^{56}$Ni yield from Type~II SNe by \citet{Rodriguez2021, Rodriguez2023} applied to multi-element abundance distributions in the Milky Way. Theoretically predicted SN yields \citep[e.g.][]{Seitenzahl2013, Sukhbold2016, Limongi2018, Gronow2021} are subject to substantial theoretical uncertainties. Empirical considerations regarding yields are often necessary for GCE models to make tenable predictions (see also discussion in \citealt{Palla2022}).

With the exception of our SN yields, we find through experimentation that the decline in \MgFe{} with time is most significantly sensitive to the value of $\tau_\star$. We therefore expect its connection to the decline in \MgFe{} on which this paper is focused to be robust. We refer to \citet{Johnson2020} and \citet{2022arXiv220402989C} for further discussion of these GCE models.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276.pdf}
  \caption{\textbf{When old stars are $\alpha$-enhanced, our subhalo of interest from TNG displays a prominent bimodality.} The upper left panel shows the distribution in the \MgFe{}-\FeH{} plane of the Milky Way, demonstrating a clear bimodality (data selection given in text). The lower left panel shows the 1D histograms of \MgFe{} at fixed \FeH{} values of $-0.5$, $-0.25$, $0$, and $0.25$ (blue, orange, green, and red, respectively). In the Milky Way, the bimodality is strongest at low metallicities while disappearing at high metallicities. The middle column shows the same plots but for our TNG subhalo of interest (392276) and with the fixed \FeH{} values $0.25\,\dex$ lower. Only faint structure is seen in the lowest bin (blue, $-0.75\,\dex$). The right column shows the same subhalo but after increasing the \MgFe{} value of star particles formed before $z=1.5$ linearly with formation time (specifically by incrementing \MgFe{} by $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$ if $t_{\textrm{form}} < t_{1.5}$, where $t_{1.5}$ is the age of the universe at $z=1.5$). A clear bimodality is shown in these panels which, unlike in the Milky Way, is present at all metallicities.}
  \label{fig:fig1}
\end{figure*}

The main result of our paper is given in Figure~\ref{fig:fig1}. Here, we compare the abundance plane in the Milky Way (left column) to that in our subhalo of interest (middle and right columns). The upper panels show the 2D distribution in the space of \MgFe{}-\FeH{}. We have applied the standard \texttt{scipy} implementation of a gaussian kernel density estimator to a Cartesian grid of points. For each panel, we normalize so that the integral of the distribution is unity. Colors are plotted in a log scale ranging from $0.08$ to $15\,\dex^{-2}$. Contour lines are plotted at $0.1$, $1.5$, and $10\,\dex^{-2}$.

The colored vertical lines are indicated at $\FeH=-0.75$, $-0.5$, $-0.25$, and $0\,\dex$ in the Milky Way, and at bins $0.25\,\dex$ higher in the simulations. The lower panels show 1D histograms of \MgFe{} in bins centered on these values. The bins for the Milky Way/simulations have width $0.2$/$0.05\,\dex$. The \MgFe{} values are given offsets so their medians are at zero. The rationale for the higher plotted \FeH{} in the simulations comes just from the empirical location of the bimodalities. The Milky Way shows a clear bimodal population, with a high-$\alpha$ sequence most clearly distinct from the low-$\alpha$ sequence at low metallicity. The two sequences merge around solar metallicity.

@@

The abundance distributions shown in Figure~\ref{fig:fig1} can be better understood by examining the evolution of \MgFe{} with time of the individual stars/star particles. In the upper panels of Figure~\ref{fig:alpha} we show the true distribution of \MgFe{} as a function of time for the fiducial SoI in the middle and for the post-processed, $\alpha$-enhanced subhalo to the right. We use age instead of formation time in order to better facilitate comparisons to observations. These panels are true 2D histograms, with a logarithmic colormap normalized to the maximum of the plot. The values of \MgFe{} are offset so their medians are at zero.

There is a gap in the ages that occurs at an age of $10.6\,\Gyr$, which we mark with a vertical dashed line. Star particles older than this line have a much clearer gradient in time with \MgFe{} than stars that form after. In the $\FeH=-0.25$ bin, star particles which form directly after this line have a slightly depressed \MgFe{} than stars which form a short time later.

In the lower panels we plot to the left data from the Milky Way and to the middle and right from the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we add $15\%$ age and $0.01\,\dex$ \MgFe{} Gaussian errors. These values are roughly in line with the observational errors from the APOKASC2 and APOGEE datasets that the lower left panel is made from (see Appendix~\ref{app:obs_err} for plots of the observational errors). The error in \MgFe{} is not signficant, but the age error, which is $1.5\,\Gyr$ at $10\,\Gyr$, smears out the distribution before and after the dashed line. The $\alpha$-enhanced SoI still shows two separate populations in this plot, but they now significantly overlap in age.

The Milky Way distribution (lower left panel) bears some resemblance to the $\alpha$-enhanced SoI. In particular, the $\FeH=-0.5$ and $-0.25$ bins (blue and orange, respectively) show what appears to be two populations overlapping in age but nonetheless separated in \MgFe{}. These are the bins at which the bimodality is strongest (upper left panel of Figure~\ref{fig:fig1}).

\subsection{Evolutionary History}\label{ssec:evol}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/392276_SFH_AGN_M200.pdf}
  \caption{\textbf{The evolutionary history of our subhalo of interest.} The upper panel shows the SFH of our subhalo. This SFH is computed using the SFR within twice the stellar half-mass radius in our SoI's main progenitors. In this and all subsequent panels we show a vertical dashed line at the same position as in Figure~\ref{fig:alpha} ($\sim3.2\,\Gyr$), which delineates the separation between the high- and low-$\alpha$ sequences. The gap in ages is naturally associated with a gap in the SFH located at the dashed line. The middle panel shows the BH accretion rate as a fraction of the maximum (Eddington) accretion rate across time. The accretion rate is high early on, but then drops. The gap in ages (dashed line) is coincident with a local peak in the BH accretion rate. In the lower panel we show the mass assembly of the halo, given as the mass within the radius that encloses $200\times$ the mean density of the universe. In this plot mergers are shown as sudden increases in $M_{200}$. Mergers at very early times when the halo is signifcantly less massive than its $z=0$ mass are not shown, but one can see that no clear merger is associated with the gap in stellar ages. Two mergers occur later on at $t\sim10$ and $\sim12\,\Gyr$.}
  \label{fig:history}
\end{figure}
